\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{booktabs}
\usepackage{url}
\usepackage[hidelinks]{hyperref}
\usepackage{xcolor}

\title{RIGOR: Research Integrity Guardrails for Open Research\\
\large An integrity-first agent for the full research workflow ---\\
grounded literature, verified experiments, honest statistics, auditable manuscripts}
\author{Alif Mahmud\\
\small Department of Electrical and Electronic Engineering\\
\small Bangladesh University of Engineering and Technology (BUET), Dhaka, Bangladesh}
\date{July 2026}

\begin{document}
\maketitle

\begin{abstract}
Large language model (LLM) agents are moving into every stage of the research
workflow---literature survey, experiment execution, statistical analysis, and
manuscript drafting---and they import three well-documented failure modes:
hallucinated citations, numbers in prose that silently drift from the data
behind them, and superiority claims that random seed variation does not
support. We present RIGOR, an open-source toolkit of eight agent
\emph{skills} (instruction files paired with small, dependency-light Python
programs) that makes agent-assisted research \emph{grounded by construction}
rather than by exhortation. RIGOR enforces three architectural guarantees:
(1)~a paper may be cited only if it exists in a provenance-tagged store
populated exclusively by live bibliographic API responses; (2)~audit tools
that reconcile bibliographies and manuscript numbers against ground truth
\emph{report and propose but never edit}; and (3)~every ``X beats Y'' claim is
gated on paired-by-seed significance tests that report exact $p$-values and
$n$, with non-significance reported as a result. Around these gates, RIGOR
automates the experiment loop itself: the agent writes and edits the
experiment code, proves every pipeline on a local smoke subset before any
GPU is spent, dispatches multi-seed sweeps to free cloud GPUs (Kaggle
unattended and recommended; Colab semi-attended and field-verified, where one
human tap doubles as the GPU
approval)---optionally scheduling across a research team's pooled per-account
quotas---and passes every completed run through an integrity checklist
before its numbers reach the researcher. The manuscript is treated the same
way: \LaTeX{} compiled locally, diagrams written as TikZ code the agent
renders and visually inspects itself, and every result table generated from
raw run artifacts---numbers flow one way, from code to prose. We describe the
design, a
two-layer portability model that separates generic mechanics from per-project
configuration, and a working case study on a real multi-seed forecasting
project in which each tool produced verifiable findings, including a wrong
bibliography title, 47 stale-drift candidates in one manuscript, and a
single-seed ``best model'' headline that dissolved under paired testing.
RIGOR is available under the MIT license.
\end{abstract}

\section{Introduction}

LLM assistance in research is no longer hypothetical: models survey fields,
draft related-work sections, orchestrate experiments, and edit manuscripts.
The efficiency gains are real, but so are the integrity risks, and they are
not hypothetical either. Controlled evaluations find that a large fraction of
LLM-generated bibliographic citations are fabricated or substantively
erroneous~\cite{walters2023fabrication}. Two further failure modes emerge as
agents move beyond writing into the experimental loop:

\begin{enumerate}
\item \textbf{Text--data drift.} Modern empirical papers regenerate their
result tables from experiment artifacts, but the surrounding prose---the
abstract, discussion sentences, figure captions---is hand-written. After a
re-run, the tables update and the prose silently keeps the old numbers. In
our case study (Section~\ref{sec:case}), a careful \emph{manual} audit of one
manuscript found nine such mismatches; the mechanized audit later surfaced
dozens of candidates.
\item \textbf{Seed-noise claims.} With $n{\approx}10$ random seeds, a nominal
gap between two models is often smaller than the seed-to-seed spread of either
one. A single-seed ranking presented as a finding is not robust, and agents
that summarize results tables reproduce such rankings uncritically.
\end{enumerate}

Prompting an agent to ``be accurate'' does not fix any of these: the failure
happens precisely when the model is confident. Our position is architectural:
\emph{integrity failures should be surfaced by mechanism, not caught by
vigilance}. RIGOR contributes a small, auditable toolkit organized around
three guarantees:

\begin{itemize}
\item \textbf{Grounded by construction} (Section~\ref{sec:lit}): every
citation must originate from a live Semantic Scholar API
response~\cite{kinney2023semanticscholar} recorded in a provenance-tagged
store; remembered papers must be verified against the API before they may be
mentioned; claims about a paper's content must trace to its stored abstract,
its full text (with page tags), or the citing sentences retrieved for it.
\item \textbf{Report, propose, never edit} (Section~\ref{sec:audit}): the
bibliography and manuscript auditors emit classified findings with suggested
fixes; a human applies each one. There is no code path by which an audit
modifies the user's files.
\item \textbf{Exact statistics} (Section~\ref{sec:stat}): superiority claims
are tested paired-by-seed (Wilcoxon signed-rank~\cite{wilcoxon1945individual}
with a paired $t$ reference~\cite{virtanen2020scipy}, optional Holm
correction~\cite{holm1979simple}); output always includes $n$ and exact
$p$-values, and a null result is reported, never dropped.
\end{itemize}

These guarantees would matter little if the agent merely \emph{advised} while
a human ran everything by hand. RIGOR therefore also automates the loop the
guarantees protect (Section~\ref{sec:auto}): the agent prepares and edits the
experiment code itself, proves it end-to-end on a local smoke subset,
dispatches unattended multi-seed sweeps to free cloud GPUs, and verifies
everything that comes back---so the integrity gates sit inside a workflow
that is genuinely hands-off between the researcher's kickoff instruction and
the returned briefing.

\section{Related work}

Commercial AI research assistants (literature-discovery products and chat
interfaces with retrieval) address parts of the citation-grounding problem,
but are generally closed-source, non-scriptable, and difficult to audit; their
guarantees, where stated, are not inspectable. On the open side,
bibliographic APIs---Semantic Scholar~\cite{kinney2023semanticscholar},
OpenAlex~\cite{priem2022openalex}, and Crossref~\cite{hendricks2020crossref}%
---provide the verifiable substrate RIGOR builds on, but are typically consumed
by ad-hoc scripts without an integrity policy connecting them to what the
agent ultimately writes. Reproducibility tooling (workflow managers,
experiment trackers) addresses artifact provenance but not the last mile where
numbers enter prose. Open manuscript-preparation systems come closest in
spirit: Manubot assembles papers from version-controlled sources with
citation-by-identifier~\cite{himmelstein2019manubot}, and showyourwork ties an
article's figures to the code that regenerates
them~\cite{showyourwork}. Both enforce reproducible \emph{manuscript assembly};
RIGOR is complementary, governing the agent-driven loop that produces the
numbers and citations \emph{before} they reach the manuscript. RIGOR's
contribution is not a new algorithm but the missing policy layer: a set of
enforceable rules, each backed by a mechanical tool, spanning literature,
bibliography, manuscript, and statistics.

\section{Design}

\subsection{Skills: judgment and mechanics, separately}

Each RIGOR capability is a \emph{skill}: a folder containing a
\texttt{SKILL.md} (the agent-facing contract: when to act, integrity rules,
how to adjudicate tool output) and---in all but the remote-execution skill,
which drives a shared runner template---one Python script (the mechanics:
API calls, parsing, classification). This split is deliberate. The scripts are
deterministic and testable offline (all HTTP is stubbed in CI); the judgment
calls---is this near-miss number a stale claim or a coincidentally similar
correlation coefficient?---are documented duties of the agent, executed under
human review. The scripts are standard-library-first Python~3.10+ (scipy and
pypdf~\cite{pypdf} only where noted) and equally usable by hand, with no agent
at all.

\subsection{Two-layer portability}

Skills contain no project paths, project names, or reference numbers
(\emph{Layer~1}). Everything project-specific lives in one profile file
(\emph{Layer~2}): manuscript path, results glob, which interpreter has scipy,
where verified reference numbers live, and safety flags. The agent reads the
profile and wires values into command-line arguments; the scripts stay
generic. Installing RIGOR into a project is a folder copy; onboarding scans
the repository, asks only what cannot be inferred, and writes the profile.
Credentials (API keys, cloud tokens) are machine-level environment variables
and never travel with the folder.

\subsection{Grounded literature work}
\label{sec:lit}

The \textbf{lit-review} skill maintains, per topic, a JSON store of papers
keyed by Semantic Scholar identifier, where every entry records
\emph{provenance}: which query, snowball, recommendation, or lookup produced
it. Sub-commands cover ranked search; exhaustive boolean \emph{bulk} search
for long-tail coverage claims; citation snowballing in both directions;
recommendations; \emph{contexts} (the sentences in which the field cites a
paper, with intent tags--- upgrading novelty screening from
regex-over-abstracts to reading actual usage); \emph{enrich} (abstracts
missing from Semantic Scholar are filled from
OpenAlex~\cite{priem2022openalex} into a separately-attributed field, never
overwriting the primary source); \emph{refresh} (batch updates of citation
counts, stamping every entry with a fetch date so ``cites $N$'' claims carry
an as-of date); validated open-access PDF retrieval (magic-byte checks with an
arXiv fallback, because publisher ``OA'' links are sometimes HTML landing
pages); and page-tagged full-text extraction, so a claim grounded in a paper's
body cites a page.

The \emph{report} command renders the collection as a document a researcher
actually reads: every paper appears as a rich entry---title, Semantic
Scholar / DOI / open-access-PDF links, citation counts, provenance, and the
best available abstract with its source attributed---and, given the
researcher's own project description as a \emph{focus}, the whole report is
re-ranked and tiered (\textsc{core} / \textsc{related} /
\textsc{peripheral}) by relevance to that focus. Relevance is a transparent
IDF-weighted term-overlap score (title weighted over summary over abstract)
whose matched terms are printed beside every paper, so the ranking is
inspectable and correctable by eye rather than a black box. The integrity
rules are short and absolute: no paper may be cited unless it is in the
store; remembered papers are verified first; content claims trace to stored
fields; citation counts are reported exactly as returned.
\textbf{topic-watch} re-runs the store's own recorded queries and diffs for
new papers, keeping a survey current before a revision.

\subsection{Auditing the bibliography and the manuscript}
\label{sec:audit}

\textbf{bib-audit} checks each BibTeX entry against Semantic Scholar and
Crossref: DOI-first, title-match fallback with similarity scoring. Verdicts
are \textsc{verified}, \textsc{mismatch} (field-level diffs and suggested
corrections), \textsc{not-found}, \textsc{non-paper-ok} (datasets, standards,
software: checked by URL liveness instead), and \textsc{unverifiable}
(submitted work). Two calibration details matter in practice. First,
bibliographic databases often record a conference paper under its preprint
year; a $\le$2-year offset is therefore reported as an informational note, not
an error---na\"ively flagging it would have produced false alarms on a third
of our case-study bibliography. Second, when a title-only match disagrees on
year by more than two years, the finding is flagged but \emph{no fix is
auto-suggested}: the match is likely a different edition or a review of the
same work, and a plausible-looking auto-fix would be worse than none.

\textbf{claims-audit} extracts every numeric literal from a LaTeX manuscript's
prose and captions (skipping generated table inputs, references, labels, and
layout dimensions) and classifies each against a ground-truth pool built from
the machine-generated tables plus the aggregated experiment results:
\textsc{matched} (within rounding), \textsc{near-miss} (close but drifted:
the prime stale-claim suspects), and \textsc{orphan} (traceable to no data
source---mostly legitimate derived or structural values, but containing the
dangerous class of results-like numbers with no provenance). It also flags
unfilled placeholder macros and any results-derived figure whose file is older
than the newest results artifact. The tool deliberately trades false positives
for zero false negatives on the dangerous class; the human pass over the
report is the point, not a failure of automation.

\subsection{Statistical honesty}
\label{sec:stat}

\textbf{stat-check} ingests per-run results files (one JSON per experiment
run), groups them into studies, deduplicates re-runs (newest per seed wins),
excludes smoke tests, and aligns metric values \emph{by seed} so tests are
properly paired. For each requested pair it reports $n$, means, the median
paired difference, the exact Wilcoxon signed-rank $p$, and a paired-$t$
reference value; a default battery tests best-vs-runner-up and best-vs-each-%
baseline. With $n{=}10$ seeds the smallest achievable two-sided Wilcoxon $p$
is $2/2^{10} \approx 0.002$; the tool documents this floor so users read
$p=0.002$ as ``a clean sweep,'' not as arbitrary precision. Holm correction is
available and its use (or absence) is disclosed in the output.

\subsection{Autonomous experimentation: write, smoke-test, dispatch, verify}
\label{sec:auto}

RIGOR's execution layer treats the agent as more than a dispatcher: the agent
\emph{writes and maintains the experiment code itself}, under the same
integrity gates as everything else. Given a training notebook, it applies the
four edits that make the notebook remotely runnable---a single tagged
parameter cell for injection~\cite{papermill}; an environment path shim so
the identical notebook runs on a laptop, on Kaggle, or on Colab; a loud
\texttt{SMOKE\_TEST} switch whose banner makes pipeline-check numbers
impossible to mistake for results; and a final cell exporting every metric,
type-cast, to a \texttt{results.json}---and then debugs its own work against
a local CPU environment. The \emph{smoke-test discipline} is a hard gate: no
configuration is dispatched to a GPU until the full pipeline has run
end-to-end on a tiny local subset, so debugging happens in seconds on the
laptop, never in minutes of donated GPU quota.

\textbf{run-remote} then drives the sweep unattended: parameter injection,
push to a free cloud GPU service, polling, artifact download, parsing, and an
append-only experiment journal. Failures are absorbed, not fatal: a failed
run is logged and retried under a fresh run identifier (so a stale journal
entry can never silently mask a retry) while the rest of the sweep proceeds.

The same layer scales to \emph{fleet dispatch}. A research group can register
each member's own cloud account and API token in a local roster (credentials
stay machine-level and never enter the repository), and the agent plans a
campaign across the pooled per-account quotas: it probes each account before
spending anything (a one-minute kernel confirming a GPU actually attaches),
stages datasets where each account needs them, partitions the sweeps, writes
one sweep file per account, launches the runners in parallel, monitors the
logs, self-heals transient failures, and returns a verified briefing rather
than a directory of raw files. The researcher's role collapses to two
moments: the kickoff instruction and the debrief. One hard rule survives
full autonomy: enabling GPU---which spends real, shared quota---requires the
owner's explicit approval, every time, in every project.

Version 1.2 adds \textbf{colab-run}, a second execution backend built around
an honest platform constraint: unlike Kaggle, free Google Colab exposes no
official headless-execution API, and RIGOR does not automate around a
provider's terms. The design is therefore \emph{semi-attended}: the agent
injects parameters (implementing the papermill injected-cell convention in
the standard library) and stages the notebook into the researcher's local
Google-Drive-synced folder, from which it appears in Colab within seconds;
execution is a single tap (\emph{Runtime}~$\to$~\emph{Run all}) in any
browser, including a phone; the notebook persists its \texttt{results.json}
back to Drive, and the agent's poller picks it up from the synced folder,
validates it, and journals it alongside the Kaggle runs. Everything but the
tap is automated---and the tap itself \emph{is} the explicit GPU approval
the safety rule requires, turning the human checkpoint from a procedural
promise into a structural necessity. The loop is \emph{field-verified}, not
merely designed: a staged notebook executed on a live Colab VM and returned
a token-stamped \texttt{results.json} through the synced folder, closing the
round trip end-to-end before the capability was claimed here
(Section~\ref{sec:case}).

\paragraph{Choosing a backend.} The two backends are not equals, and the
toolkit says so plainly. \textbf{Kaggle is the recommended default}: its
official kernels API makes the entire loop headless---push, poll,
download---which is what enables true fleet dispatch, overnight campaigns,
and the two-moment (kickoff/debrief) workflow. \textbf{colab-run is the
deliberate fallback}: one tap per run automates everything \emph{around}
execution but not execution itself, so a ten-run sweep means ten taps.
Reach for Colab when the week's Kaggle quota is exhausted, when a
Colab-specific runtime is wanted, or when the per-run human checkpoint is
itself desirable; route everything else through the Kaggle path.

\textbf{verify-run} is the checklist every completed run passes before its
numbers reach the researcher: all expected configurations present (a missing
key is a silent mid-run failure, not a partial success); smoke-test flags
honored (pipeline-check numbers are never reported as results); early-abort
(futility) stops disclosed rather than hidden; NaNs reported as failures;
parameter counts matched exactly against anchors (an exact match fingerprints
that the intended pipeline ran); and seed counts always disclosed, so a
single-seed number is never presented as a robust average.

\subsection{The manuscript layer: local \LaTeX{}, agent-drawn figures, one-way numbers}
\label{sec:paper}

RIGOR's workflow keeps the entire manuscript---text, tables, and diagrams---as
plain text that compiles \emph{locally}, with the agent as typesetter and
illustrator. This deliberately displaces three tools researchers usually
reach for:

\begin{itemize}
\item \textbf{Overleaf} is replaced by a local \texttt{latexmk} build inside
the researcher's editor: no compile-minute limits, no internet dependence,
the full paper rebuilding in seconds, and the manuscript versioned in Git
alongside the code that produced its numbers.
\item \textbf{Diagramming tools (draw.io and kin)} are replaced by TikZ
\emph{written by the agent}: an architecture diagram is code, so the
researcher describes the figure in plain English and the agent writes the
drawing---an infinitely-zoomable vector figure that remains editable, by the
agent, for the life of the paper, instead of a mouse-drawn artifact that
must be redrawn by hand at every revision.
\item \textbf{AI image generators} are displaced for schematic figures by the
same mechanism: TikZ output is exact, editable, and diff-able, where
generated raster images are approximate, frozen, and blurry at print
resolution---and subject to usage caps besides.
\end{itemize}

What makes this workable in practice is that the agent \emph{closes its own
visual quality loop}: after compiling a figure it rasterizes the PDF, reads
the image back, spots overlapping labels, clipped boxes, or bad spacing,
edits the TikZ source, and recompiles---iterating until the figure is clean.
A loop that previously required a human, a mouse, and an afternoon runs
without either.

Result tables and result figures never pass through human hands at all: an
aggregator script reads every \texttt{results.json}, computes
mean~$\pm$~std across seeds, and writes \texttt{\textbackslash input}-ready
table fragments and figure PDFs directly into the manuscript tree. Numbers
therefore flow \emph{code $\rightarrow$ results.json $\rightarrow$
aggregator $\rightarrow$ generated tables and figures}, in one direction
only: no metric is ever retyped by hand into prose, and if a number looks
wrong the fix belongs in the code or the data, never in the text.
claims-audit (Section~\ref{sec:audit}) exists precisely to catch the places
where that rule was broken by hand before the pipeline enforced it.

\section{Case study: a real project as first install}
\label{sec:case}

RIGOR's first install is a live research project---a multi-seed regional
electricity-load forecasting study with a LaTeX manuscript, a 14-entry
bibliography, a 245-paper literature collection, and six 10-seed experiment
studies executed remotely. Each tool produced concrete, verifiable findings:

\begin{itemize}
\item \textbf{bib-audit} (14 entries, 51\,s of wall time): one reference's title turned out to be
a paraphrase of the real paper's title (the entry resolved to nothing until
the audit surfaced the correctly-titled record); three title-only matches hit
different editions or records and were flagged with no auto-fix; two dataset
resources were verified by URL liveness; five missing DOIs were recovered.
The preprint-year calibration prevented three false alarms.
\item \textbf{claims-audit} (one manuscript): 271 numeric claims classified
(151 matched, 47 near-miss, 73 orphan) in 0.25\,s---mechanizing, and
extending, a manual audit that had found nine real text--data mismatches, and
leaving the human only the semantic adjudication of the flagged candidates.
\item \textbf{stat-check} (14 models $\times$ 10 seeds): the nominally best
model was statistically indistinguishable from several siblings (Wilcoxon
$p\approx0.5$), while every family-versus-baseline comparison survived exactly
($p=0.002$--$0.02$). A single-seed ``best architecture'' headline from an
earlier phase of the project did not survive; the published claim was reshaped
accordingly---the tool's purpose working as intended, against its own
project's preferred narrative.
\item \textbf{run-remote / verify-run}: the six 10-seed studies were executed
as unattended GPU sweeps---more than sixty runs plus retries, dispatched by
the agent across multiple registered free-tier accounts, every run passing
the verification checklist before aggregation. The pre-flight probe caught
one account whose GPU access was silently unavailable and rebalanced the
schedule onto the rest; a transient network outage mid-campaign was healed by
fresh-identifier retries, with no human intervention between kickoff and
briefing.
\item \textbf{colab-run} (field verification): the Colab loop was exercised
live before being claimed in this paper---a staged notebook executed on a
Colab VM and returned a token-stamped \texttt{results.json} through the
synced folder, matching the injected token and seed exactly. The field test
immediately exposed a real parameter-injection bug the unit tests had
missed: Python-style \texttt{True}/\texttt{False} values were being parsed
as JSON, injecting the truthy \emph{string} \texttt{'False'}, which would
have made smoke-test mode impossible to disable. It was fixed and locked in
as a regression test. We publish no capability we have not exercised---the
same standard the toolkit enforces on its users' claims.
\item \textbf{lit-review}: the collection was built from seven ranked queries,
snowballing, recommendations, and a boolean bulk pass; abstract enrichment
filled 9 of 39 missing abstracts from OpenAlex (the remainder are withheld by
publishers---an honest, documented gap); a single batch refresh re-dated all
citation counts. With the project's own description as the report's
\emph{focus}, the 245-paper collection tiers into 11 \textsc{core} / 95
\textsc{related} / 139 \textsc{peripheral}---and the \textsc{core} tier
surfaced, on its own, the same closest-rival paper a manual screen had
previously identified.
\item \textbf{manuscript layer}: all five of the case-study paper's
architecture and concept diagrams are agent-written TikZ, iterated under the
agent's own render-and-inspect loop; every result table and figure is
aggregator-generated from the raw run artifacts. Since the one-way pipeline
was adopted, no metric has been hand-typed into that manuscript's prose.
\item This paper's own bibliography was verified by resolving every DOI
through doi.org content negotiation before inclusion.
\end{itemize}

Nothing in the scripts is specific to the case study's domain. To make that
concrete---and reviewable without any API key, GPU, or agent
harness---the repository ships fully offline, synthetic worked examples
(\texttt{examples/}) for the statistics and audit tools, plus a network demo
in which \textbf{bib-audit} catches a deliberately fabricated citation; the
offline examples are exercised end-to-end by the CI test suite, so the
documented walkthroughs cannot drift from the code.

\section{Limitations}

RIGOR's grounding is only as complete as its substrate: Semantic Scholar and
OpenAlex have coverage gaps (some publishers withhold abstracts; very recent
or non-English work may be missing), so ``not found'' is evidence, not proof,
of non-existence. Rate limits (1~request/s on the primary API) make large
collections slow to build and briefly throttle bursty use.
The claims auditor is calibrated for LaTeX manuscripts with generated tables;
purely hand-written papers reduce it to placeholder and figure checks. The
statistical tool assumes per-seed results files in a simple JSON schema;
projects must export it (a few lines in the training notebook). Paired tests
at $n{\approx}10$ have limited power, which the tool discloses rather than
solves. Finally, the skill layer's semantic adjudication depends on the
hosting agent following written rules; the guarantees that matter most
(no citation outside the store, no auto-edits, exact $p$-values) are enforced
by the tools themselves, but the quality of, e.g., near-miss adjudication
varies with the agent.

\section{Conclusion}

RIGOR treats research integrity as an architecture problem: identify the
places where AI assistance can silently fabricate, drift, or overclaim, and
put a mechanical, auditable gate at each one. The result is a small toolkit
that a researcher can copy into any project, drive with or without an agent,
and extend---with the guarantee that everything it lets through leaves a
verifiable trail.

\paragraph{Future work.} The execution layer is backend-agnostic by design
(prepare, stage, execute, collect, verify), and v1.2's Colab backend already
lets groups pool quota across two free services. Next: fully headless Colab
execution through the paid Colab Enterprise API for funded groups, scheduled
topic watching for teams with always-on machines, and a multi-user
deployment study of RIGOR across a working research group.

\paragraph{Availability.} RIGOR is available under the MIT license at
\url{https://github.com/alif199339/rigor}. Collections built with it
rely on the Semantic Scholar API; published material based on them should cite
Kinney et al.~\cite{kinney2023semanticscholar} per the API license.

\bibliographystyle{unsrt}
\bibliography{refs}

\end{document}
